---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

We load in the activity.csv file and can see from the stucture of the rawdata that there are 3 columns of data: $steps, $date, $interval.

```{r}
      activity <- read.csv("./activity.csv") 

      str(activity)
```

We use the dplyr library to group_by the dates and sum the total number of steps

```{r}
      library(dplyr)
      
      by_date <- group_by(activity, date)
      
      per_day <- summarise(by_date,
            total_steps = sum(steps, na.rm = TRUE))
```


## What is mean total number of steps taken per day?


We can draw a histogram to show the distribution of the totals for each date and calculate the mean and median total_steps.

```{r}
      hist(per_day$total_steps)
      
      steps_mean <- mean(per_day$total_steps)
      steps_median <- median(per_day$total_steps)
      
      print(steps_mean)
      print(steps_median)
      
```


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
